---
title: "README"
author: "Thomas"
author: "Sean Fahey"
date: "3/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Rmonkey** provides access to [Survey Monkey](https://www.surveymonkey.com/), for the complete integration of survey data collection and analysis into a single, easily reproducible workflow.

## Installation ##

This version of **Rmonkey** is updated to interface with the SurveyMonkey API v3.  The latest development version, available here, can be installed directly using  [devtools](http://cran.r-project.org/web/packages/devtools/index.html):

```{r}
if(!require("devtools")) {
  install.packages("devtools")
  library("devtools")
}
install_github("seanofahey/Rmonkey")
library("Rmonkey")
```

## Setup ##

To use Rmonkey, the user must first have a Survey Monkey account which can be obtained at https://www.surveymonkey.com/user/sign-in/.  Next, within the SurveyMonkey account, the user must create an app that can be used to access data via the api.  This can be done at https://developer.surveymonkey.com/apps/.  In the app configuration, set OAuth Redirect URL as `http://localhost:1410`.  Finally set the user permissions for creating and viewing data within SurveyMonkey with the scopes options.

Once the app is configured, click on settings to reveal the Client ID and Secret keys.  These can be loaded into R using `options`:

```{r}
options(sm_client_id = 'YourClientID')
options(sm_secret = 'YourAPISecret')
```

Rmonkey uses these values inside `smlogin` to initiate an OAuth2.0 login. Calling `smlogin()`, you will redirected to your web browser, where you will login with your regular Survey Monkey account information. `sm_login` will then store a durable OAuth token in `options('sm_oauth_token')`, which is automatically retrieved in subsequent Rmonkey operations.

This token is currently long-lived (meaning it is valid indefinitely). This means that saving the OAuth token between R sessions will prevent you from having to login each time you load **Rmonkey** and allow you to use the package in non-interactive R sessions. If you have trouble logging in, it is also possible to copy the OAuth access token from the [App Settings](https://developer.surveymonkey.com/apps), which can then be manually stored in `options('sm_oauth_token')`. 

## Code Examples ##

Below are some code examples showing how to use the package.

### Establish and Test the API Connection ###

To establish a connection between R and SurveyMonkey use the `smlogin()` function.  This will open an interactive session in your browser to present the API permissions and request authorization.  This function completes the OAuth handshake and saves a long lasting token on the computer.

```{r}
smlogin()
```

To verify that the connection is functional you can retrieve information about the user with the `userdetails()` function.

```{r}
userdetails()
```

## Get a list of Surveys ###

**RMonkey** provides several options for retrieving information about the surveys in the account.

One can retrieve a list of surveys using the `surveylist()` command.  This will return a list with details of each survey.

```{r}
sl <- surveylist()
head(sl)
```

To retrieve a list of surveys that have been modified since a certain date one can use the `start_modified_at` parameter within the `surveylist()` function.

```{r}
sl <- surveylist(start_modified_at = '2017-03-25')
head(sl)
```

Additional parameters can be used to change the number of responses, add fields to the survey list response, and sort the responses.

```{r}
sl <- surveylist(per_page = 100, include = 'response_count', sort_by = 'num_responses', sort_order = 'desc')
head(sl)
```

## Get Details about a survey ##

To see details about a single survey use the `surveydetails()` function.  This will return basic information about the survey including the title, nickname, ID, number of questions, number of respondents, etc... 

```{r}
s1.d <- surveydetails(sl[[1]])
s1.d
```

## Preview a Survey in the Browser ##

To see a preview of a survey use the `surveypreview()` function.  In the function, pass a survey object retrieved using the survey list function.  This will open a tab in your browser to display the survey preview.  

```{r}
surveypreview(sl[[1]])
```

## Retrieve Survey Responses ##

To get a list of responses for a survey use the `surveyresponses()` function.  In the function, pass a survey object retrieved using the survey list function.  This will return a data frame with one row per response and one column per question.  (NOTE: This can take a long time to run.)

```{r}
s1.r <- surveyresponses(sl[[5]])
head(s1.r)
```

To get the survey results into a columnar format use the response_format = 'column' parameter.  (This can be useful if exporting the data to systems that ingest data in this format.)

```{r}
s1.r <- surveyresponses(sl[[5]], response_format = 'column')
head(s1.r)
```

